Concurrent Programming with CXL

### 2020-06-05 book/paper.tex
Working on semaphore section

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{csbarebones.cxl}
\end{code}
\caption{A barebones critical section.}
\label{fig:csbarebones}
\end{figure}

@@

A good example is the so-called \emph{bounded buffer}.  It is essentially
a queue implemented using a circular buffer of a certain length and two pointers:
one where items are inserted and one from where items are extracted.  If the
buffer is full, processes that want to add more items have to wait.  In the usual
formulation, processes that want to extract an item when the buffer is empty
also have to wait.  It would be easy to build such a thing using a busy waiting
approach, but again, we would like to avoid busy waiting.

@@

>>>>

\begin{figure}
\begin{code}
\VerbatimInput[xleftmargin=5mm,numbers=left]{csbarebones.cxl}
\end{code}
\caption{A barebones critical section with two processes.}
\label{fig:csbarebones}
\end{figure}

@@

A good example is the so-called \emph{bounded buffer}.  It is essentially
a queue implemented using a circular buffer of a certain length and two pointers:
one where items are inserted and one from where items are extracted.  If the
buffer is full, processes that want to add more items have to wait.  In the usual
formulation, processes that want to extract an item when the buffer is empty
also have to wait.
This problem is known as the ``Producer/Consumer Problem.''
It would be easy to solve using a busy waiting
approach, but again, we would like to avoid busy waiting.

@@

One can think of a semaphore as a generalization of a lock.  In fact, a
lock can be implemented by a semaphore initialized to 1.  Then \texttt{P}
acquires the lock, and \texttt{V} releases the lock.

Figure~\ref{fig:boundedbuffer} presents the implementation of a bounded
buffer using semaphores.  It features a circular bounded buffer \texttt{buf} with
slots numbered 1 through \texttt{NSLOTS}.  There are two indexes into
the buffer: \texttt{b\_in} specifies where the next produced item is inserted,
while \texttt{b\_out} specifies from where the next can be consumed.
As there may be multiple producers and multiple
consumers, updates to the indexes, which are shared variables, must be protected.
To this end we use two semaphores as locks: \texttt{l\_in} for \texttt{b\_in}
and \texttt{l\_out} for \texttt{b\_out}.

Without additional synchronization, the indexes may overflow and point to invalid
entries in the circular buffer.
To this end there is a semaphore \texttt{n\_full} that
keeps track of the number of filled entries in the buffer and a semaphore
\texttt{n\_empty} that keeps track of the number of empty entries in the buffer.

To add an item to the bounded buffer (\texttt{produce(item)}), the producing
process first has to wait until there is room in the bounded buffer.
To this end, it invokes \texttt{P(n\_empty)}, which waits until
\texttt{n\_empty > 0} and atomically decrements \texttt{n\_empty} once this
is the case.  Next, it adds the item to the next position in the buffer.
Since there may be more than one empty slot, multiple producers may attempt
to do the same thing concurrently.  Hence the use of the \texttt{l\_in}
semaphore used as a lock.  Finally, once the item is added, the
\texttt{n\_full} semaphore is incremented to signal to consumers that
an item is available.

The consumer code is symmetric.

Note the quite different usage of the \texttt{l\_} semaphores and \texttt{n\_}
semaphores.
People often call semaphores that are used as locks \emph{binary semaphores},
as their counters only take on the values 0 and 1.
They protect critical sections.
The second type of semaphores are called \emph{counting semaphores}.
They are essentially used to send signals between processes.
However, binary and counting semaphores are implemented the same way.
